Interactive Data Exploration with Smart Drill-Down (Extended Version)

### 2016-12-04 TableSummarization/journalVersion/smartDrillDown.tex
combined setting parameters with Extensions

\subsection{NP-Hardness for Problem~\ref{prob:optimal-rule-set}}
We reduce the well known {\sc NP-Hard} {\em Maximum Coverage Problem} (MCP) to a special case of Problem~\ref{prob:optimal-rule-set};
thus demonstrating the {\sc NP-Hard}ness of Problem~\ref{prob:optimal-rule-set}. MCP is given below: 
\begin{problem}\label{prob:maximum-coverage}
Given a universe set $U$, an integer $k$, and a set $S = \left\lbrace S_1, S_2, ... S_m \right\rbrace$ of subsets of $U$ (so each $S_i \subset U$), find $S^{\prime} \subset S$ such that $|S^{\prime}| = k$, which maximizes $\text{Coverage}(S^{\prime}) = |\bigcup_{s \in S^{\prime}} s|$.
\end{problem}
Thus, the goal of MCP is to find a set of $k$ of the given subsets of $U$ whose union `covers' as much of $U$ as possible. We can reduce an instance of MCP (with parameters $U, k, S$) to an instance of Problem~\ref{prob:optimal-rule-set}, which gives us the following lemma:
\begin{lemma}
Problem~\ref{prob:optimal-rule-set} is {\sc NP-Hard}.
\end{lemma}
%%%%%%%%%%%%%%%%%%%%
\begin{proof}
Consider a table with $|U|$ rows (one per element of $U$) and $m$ columns (one per $S_j$, that has value $1$ in row $i$, column $j$ if the $i^{th}$ belongs to $S_j$ and $0$ otherwise. And consider a weighting function with $W(r) = 1$ if there is at least one $1$ in $r$, and $0$ otherwise. Let $k \leq m$.

Then a rule with multiple $1$s is clearly dominated (or matched) by a subrule with only one $1$. Moreover, a rule-list that has two rules with a $1$ in the same column is dominated by one that that $1$s in different columns. Then the score of a rule list that has rules with $1$s in column set $C \subset \{1, 2,\ldots m\}$ with $|C| = k$ has score equal to the size of union $\bigcup_{c \in C} S_c$. Thus maximizing score is equivalent to maximizing the size of the union of $k$ sets, which is the Maximum Coverage Problem. Thus the NP-Hardness of Maximum Coverage implies that Score maximization is NP-hard.
\end{proof}
%%%%%%%%%%%%%%%%%%%%

@@

The parameter $m_w$ stands for \textit{Max Weight}. This parameter tells the algorithm to assume that all rules that appear in the optimal solution are going to have weight $\leq m_w$. Thus, if $S_o$ denotes set of rules with maximum score, then as long as $m_w \geq \textrm{max}_{r \in S_o}W(r)$, BRS is guaranteed to return $S_o$. On the other hand if $m_w < W(r)$ for some $r \in S_o$, then there is a chance that the set returned by BRS does not contain $r$. BRS runs faster for smaller values of $m_w$, and may only return a suboptimal result if $m_w < \textrm{max}_{r \in S_o}W(r)$. In practice, $\textrm{max}_{r \in S_o}W(r)$ is usually small. This is because as the size (and weight) of a rule increases, its Count falls rapidly. The Count tends to decrease exponentially with rule size, while Weight increases linearly for common weight functions (such as Size). Thus, rules with high weight and size have very low count, and are unlikely to occur in the optimal solution set $S_o$. Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set tend to be small. \papertext{In the technical report~\cite{DBLP:journals/pvldb/JoglekarGP15}, we describe strategies for setting $k$, $m_w$ and other parameters.} \techreporttext{Later in this section, we describe strategies for setting $m_w$ as well as other parameters.}

% TODO: scheme/heuristic for finding good m_w values, since it seems arbitrary. Later, experiments testing effect of m_w on running time, and experiments to test finding heuristic, and show that values are in fact small.

@@

%%%%%%%%%%%%%%%%%%%%%%%
\textbf{Runtime analysis:}
Our algorithm runs calls the function for finding best marginal rule $k$ times. Thus the total computational cost is about $k$ times the cost of computing the best marginal rule. Let $c_n$ be the number of counts in $C_n$ at the end of the $n^{th}$ iteration (that is, after eliminating rules that have no chance of having the highest marginal value). Also, let $c_0 = 1$.

The cost of the finding the best marginal rule can be split into three parts: (i) Enumerating all new rules to add to $C_n$ at the start of the $n^{th}$ iteration. (ii) Determining which of the rules in $C_n$ to eliminate (iii) Finding counts and marginal values for the remaining $c_n$ rules in $C_n$. The number of super-rules we can consider for (i) can be no larger than $c_{n-1}|T|$, since each rule in $C_{n-1}$ can give us at most $|T|$ new candidate super-rules for $C_n$. The cost of (ii) is $O(|C_n|)$. Finally, the cost of (iii) is again $c_n|T|$, since we have two nested for loops over $T$ and $C_n$ respectively. Thus our total cost is $O(|T|\sum_i c_i)$.

In the worst case, we may need to evaluate all possible rules; since the total possible number of rules is $|T||C|$ where $C$ is the set of columns, our cost is at most $O(k|C||T|^2))$. However, note that this worst case is unlikely to occur when we use the Size or Bits weighting function, because the count of a rule decreases rapidly with its size, making it unlikely that we will need to find counts for its super-rules.

In fact, if we assume that values in columns occur independently of each other, and that the most frequent value occurs $x < 1$ fraction of the time, then the number of candidate rules is bounded by $c_i \leq c_1 x^i$. This gives us $\sum_i c_i \leq c_1\sum_i x^i \leq c_1/(1-x)$.
Thus our cost is now only $O(k|T|c_1/(1-x))$ for some constant $x$. If the number of distinct values in the table ($c_1$), $k$ and $x$ are taken to be constants (i.e. not scale with table size), then our cost is $O(|T|)$.

\textbf{Approximation ratio:}
If our $m_w$ is higher than the actual weight of every rule in the best rule list, then the only approximation ratio we incur is that in the greedy approximation, which is a multiplicative factor of up to $1 - 1/e$. On the other hand, if our $m_w$ estimate is smaller than the actual maximum weight $m_w^*$ of a rule in the best list, then we may end up picking a sub-optimal rule in the best marginal rule function.

We now bound the sub-optimality of our chosen rule. Suppose the best rule to pick is $r^*$, and we pick rule $r$. Since we estimated $r^*$'s weight to be at most $m_w$, we would have estimated $r^*$ to have marginal value $\text{Est}(r^*) = \sum_{t \in r^*}(m_w - w(t))$ where $t \in r^*$ iterates over all tuples covered by $r^*$, and $w(t)$ is the weight of the current highest weight rule covering $t$. Since we picked $r$ over $r^*$, it means $\text{MarginalValue}(r) \geq \text{Est}(r^*)$. Suppose $m$ is the highest weight of a rule selected in the previous steps of BRS (so $w(t) \leq m$ for all $t$). Then the actual marginal value of $r^*$ is $\leq \frac{m_w - m}{m_w^* - m}\text{Est}(r^*)$. Thus our pick of $r$ is at most a $\frac{m_w - m}{m_w^* - m}$ factor worse than the optimal pick. If we always choose  our $m_w$ such we have $\frac{m_w - m}{m_w^* - m} \geq w$, for some constant $w < 1$, then our final score is guaranteed to be withing a $w(1-1/e)$ factor of the optimal score.
%%%%%%%%%%%%%%%%%%%%%%%

\techreporttext{
\subsection{Setting parameters $W$, $k$, $m_w$}\label{sec:setting-paramters}
Our system allows the user to tune the smart drill-down by adjusting a number of parameters. Having a lot of tunable parameters can increase the difficulty of using a system by increasing the decision-making burden on the user. To counteract this, we now provide ways to guide the user while selecting appropriate parameter values.

The $m_w$ parameter lets the user trade off the accuracy of the optimal rule-list and the running time. Ideally we want $m_w$ to equal the actual maximum weight of a rule in the optimal rule-list; this way we get full accuracy while also optimizing run-time. We cannot know the ideal value of $m_w$ in advance, but we can easily estimate it using sampling. We create a small random sample of tuples from the table, and run the BRS algorithm on it. Then the maximum weight $x$ of the output on the sample is likely to equal the maximum weight of the actual output. To account for sampling error, we can set $m_w$ to $2x$, which works well in practice.
}

%%%%%%%%%%%%%%%%%%
While the setting of $W$ (and hence $m_w$ is mostly specific to the user, we will consider a simple model of how frequently values occur in a table and determine a weighting function to favor rules of a particular size based on that model.

Let us consider a parametric weight function which generalizes the `Bits Weighting Function' from Section~\ref{sec:weighting}. 
$W(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^k$, where the $w_c$'s and $k$ are parameters of the weighting function, and $o_{r,c}$ ($o$ stands for occurrence) is $1$ if $r(c) \neq \star$ and $0$ otherwise. The Size and Bits weighting functions can be seen as special cases of the above function with $k=1$ and with $w_c$ equal to $1$ for Size and $\text{log}(|c|)$ for Bits.

We want to estimate the Score contribution of a rule according to this weight function, which requires estimating the count of the rule. Suppose for each column $c$, the most frequent value occurs $f_c$ fraction of the time. If the probability of a value occurring in a column of a tuple is independent of other values in the tuple, the $Count$ of rule $r$ is approximately $|T|\Pi_{c \in C}f_c^{o_{r,c}}$. Thus the score of a rule set containing $r$ alone can be approximated by:
$$S(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^k|T|\Pi_{c \in C}f_c^{o_{r,c}}$$

Now consider the rule that has the highest weight times count according to the above weighting function. We can approximate this rule by relaxing $o_{r,c}$ to a real number; that is by maximizing the estimated $S(r)$ subject to $0 \leq o_{r,c} \leq 1$. According to the KKT conditions, at the optimum, the partial derivative of $W(r)$ with respect to $o_{r,c}$ must be $\leq 0$ for each $c$ where $o_{r,c} = 0$, equal to $0$ if $0 < o_{r,c} < 1$ and $\geq 0$ otherwise. For each $c$, we have:
$$\frac{\partial S(r)}{\partial o_{r,c}} = S(r)(\frac{kw_c}{\sum_{c \in C} o_{r,c}w_c} + \ln f_c)$$

Thus if $x = \frac{k}{\sum_{c \in C} o_{r,c}w_c}$, then the maximum scoring rule should instantiate columns where $\frac{\ln f_c}{w_c} > x$ and some of the ones with $\frac{\ln f_c}{w_c} = x$. Thus for Size weighting, the highest score rule should select columns with high values of $\ln f_c$ i.e. highest frequency values. In contrast, Bits will select columns with highest values of $\log_{|c|}(f_c)$ i.e. values that are frequent relative to the number of distinct values in the column.

If we want the score to treat all columns equally, then the derivative must be $0$ for all $c$. This implies that for all $c$, we must have $w_c \propto \ln f_c$. Notice that Bits satisfies the above criterion if we assume that values in each column are uniformly distributed (that is, $f_c = 1/|c|$). This provides an intuitive justification for we might want to use the Bits weighting function.

We can also estimate the size of the optimal rule under the above weighting function. Specifically, the weighted fraction of columns with non-$\star$ values is given by $\frac{\sum_{c\in C} o_{r,c}w_c}{\sum_{c \in C} w_c}$ which equals $-k/(\sum_{c \in C} \ln f_c)$. Thus if we wanted our highest score rule to have $s$ fraction of the columns get instantiated, we can achieve that by setting $k = -s\sum_{c \in C} \ln f_c$.

The above reasoning also lets us estimate the $m_w$ parameter we need to use. Specifically, the highest score rule above has weight $(\frac{-\ln f_c}{kw_c})^k$.

We can use the above estimate to guess the size of the max scoring rule for the Bits weighting function. But note that in practice, values are not distributed uniformly (so $f_c$ is actually higher than $1/|c|$) and are not independent of each other, due to correlations. As a result, our estimate of the maximum size is a significant under-estimate. In our experiments (Section~\ref{sec:exp_qualitative}), we observe that use the Bits weighting function actually gives us rules of size $3$ on our datasets, while our estimate above is approximately $0.74$. As expected, the values displayed by the rules are ones that occur much more frequently relative to other values in that column.

%This evaluates to $1.44$, while the observed value in our experiments is $10$. 
%%%%%%%%%%%%%%%%%%

@@

\techreporttext{
\vspace{-5pt}
\section{Extensions}\label{sec:extensions}
\subsection{Dealing with Numerical Attributes}\label{sec:extensions-numerical}
Our framework assumes that all attributes are categorical. Attributes that have a large domain tend to have fewer tuples per value, and hence don't appear in rule summaries. Thus our algorithm does not summarize information about numerical attributes.

However, we can modify the algorithm to deal with numerical attributes. Suppose we have a numerical attribute $A$. We can create buckets for values of $A$. We choose a number of buckets $b$, and divide the range of values of $A$ into $b$ intervals, with one bucket per interval. We can create buckets having equal range sizes, or decide their range such that there is an approximately equal number of tuples in each bucket. Then we run BRS, treating the bucket id as a categorical attribute. This is already done in our MD dataset, where numerical attributes like age are divided into buckets ($18-24$, $25-34$ and so on).

\vspace{-5pt}
\subsection{Using Sum instead of Count}\label{sec:extensions-sum}
Throughout the paper, we define the total score of a rule-list using the marginal counts of rules in the list, and display the count of each rule in our table summary. However, if we have a numerical column (i.e. a `measure' column) in the table, it is straightforward to extend our summary to the `Sum' aggregate over that column instead. Suppose we are given a measure column $c_m$. Then the {\em Sum} for a rule can be defined to be the sum of $c_m$ values over all tuples covered by the rule. {\em MSum} of a rule $r$ in a rule-list $R$ is the sum of $c_m$ values over all tuples covered by $r$ and not covered by any rule in $R$ that occurs before $r$. The Score for $R$ becomes $\text{Score}(R) = \sum_{r\in R} MSum(r,R)W(r)$. Algorithm~\ref{algo:best-rule-set} can be modified to find the best rule set using the new definition of Score, simply by replacing $Count(r)$ by $Sum(r)$ and computing sum and marginal sum instead of count and marginal count in each pass over the table.
}

>>>>

\subsection{NP-Hardness for Problem~\ref{prob:optimal-rule-set}}
We reduce the well known {\sc NP-Hard} {\em Maximum Coverage Problem} (MCP) to a special case of Problem~\ref{prob:optimal-rule-set};
thus demonstrating the {\sc NP-Hard}ness of Problem~\ref{prob:optimal-rule-set}. MCP is given below: 
\begin{problem}\label{prob:maximum-coverage}
Given a universe set $U$, an integer $k$, and a set $S = \left\lbrace S_1, S_2, ... S_m \right\rbrace$ of subsets of $U$ (so each $S_i \subset U$), find $S^{\prime} \subset S$ such that $|S^{\prime}| = k$, which maximizes $\text{Coverage}(S^{\prime}) = |\bigcup_{s \in S^{\prime}} s|$.
\end{problem}
Thus, the goal of MCP is to find a set of $k$ of the given subsets of $U$ whose union `covers' as much of $U$ as possible. We can reduce an instance of MCP (with parameters $U, k, S$) to an instance of Problem~\ref{prob:optimal-rule-set}, which gives us the following lemma:
\begin{lemma}
Problem~\ref{prob:optimal-rule-set} is {\sc NP-Hard}.
\end{lemma}
\begin{proof}
\mrj{Consider a table with $|U|$ rows (one per element of $U$) and $m$ columns (one per $S_j$, that has value $1$ in row $i$, column $j$ if the $i^{th}$ belongs to $S_j$ and $0$ otherwise. And consider a weighting function with $W(r) = 1$ if there is at least one $1$ in $r$, and $0$ otherwise. Let $k \leq m$.}

\mrj{Then a rule with multiple $1$s is clearly dominated (or matched) by a subrule with only one $1$. Moreover, a rule-list that has two rules with a $1$ in the same column is dominated by one that that $1$s in different columns. Then the score of a rule list that has rules with $1$s in column set $C \subset \{1, 2,\ldots m\}$ with $|C| = k$ has score equal to the size of union $\bigcup_{c \in C} S_c$. Thus maximizing score is equivalent to maximizing the size of the union of $k$ sets, which is the Maximum Coverage Problem. Thus the NP-Hardness of Maximum Coverage implies that Score maximization is NP-hard.}
\end{proof}

@@

The parameter $m_w$ stands for \textit{Max Weight}. This parameter tells the algorithm to assume that all rules that appear in the optimal solution are going to have weight $\leq m_w$. Thus, if $S_o$ denotes set of rules with maximum score, then as long as $m_w \geq \textrm{max}_{r \in S_o}W(r)$, BRS is guaranteed to return $S_o$. On the other hand if $m_w < W(r)$ for some $r \in S_o$, then there is a chance that the set returned by BRS does not contain $r$. BRS runs faster for smaller values of $m_w$, and may only return a suboptimal result if $m_w < \textrm{max}_{r \in S_o}W(r)$. In practice, $\textrm{max}_{r \in S_o}W(r)$ is usually small. This is because as the size (and weight) of a rule increases, its Count falls rapidly. The Count tends to decrease exponentially with rule size, while Weight increases linearly for common weight functions (such as Size). Thus, rules with high weight and size have very low count, and are unlikely to occur in the optimal solution set $S_o$. Our experiments in Section~\ref{sec:experiments} also show that the weights of rules in the optimal set tend to be small. \papertext{In the Extensions Section of the technical report~\cite{DBLP:journals/pvldb/JoglekarGP15}, we describe strategies for setting $k$, $m_w$ and other parameters.} \techreporttext{Later in the extensions section, we describe strategies for setting $m_w$ as well as other parameters.}

@@

\textbf{Runtime analysis:}
\mrj{Our algorithm runs calls the function for finding best marginal rule $k$ times. Thus the total computational cost is about $k$ times the cost of computing the best marginal rule. Let $c_n$ be the number of counts in $C_n$ at the end of the $n^{th}$ iteration (that is, after eliminating rules that have no chance of having the highest marginal value). Also, let $c_0 = 1$.}

\mrj{The cost of the finding the best marginal rule can be split into three parts: (i) Enumerating all new rules to add to $C_n$ at the start of the $n^{th}$ iteration. (ii) Determining which of the rules in $C_n$ to eliminate (iii) Finding counts and marginal values for the remaining $c_n$ rules in $C_n$. The number of super-rules we can consider for (i) can be no larger than $c_{n-1}|T|$, since each rule in $C_{n-1}$ can give us at most $|T|$ new candidate super-rules for $C_n$. The cost of (ii) is $O(|C_n|)$. Finally, the cost of (iii) is again $c_n|T|$, since we have two nested for loops over $T$ and $C_n$ respectively. Thus our total cost is $O(|T|\sum_i c_i)$.}

\mrj{In the worst case, we may need to evaluate all possible rules; since the total possible number of rules is $|T||C|$ where $C$ is the set of columns, our cost is at most $O(k|C||T|^2))$. However, note that this worst case is unlikely to occur when we use the Size or Bits weighting function, because the count of a rule decreases rapidly with its size, making it unlikely that we will need to find counts for its super-rules.}

\mrj{In fact, if we assume that values in columns occur independently of each other, and that the most frequent value occurs $x < 1$ fraction of the time, then the number of candidate rules is bounded by $c_i \leq c_1 x^i$. This gives us $\sum_i c_i \leq c_1\sum_i x^i \leq c_1/(1-x)$.
Thus our cost is now only $O(k|T|c_1/(1-x))$ for some constant $x$. If the number of distinct values in the table ($c_1$), $k$ and $x$ are taken to be constants (i.e. not scale with table size), then our cost is $O(|T|)$.}

\textbf{Approximation ratio:}
\mrj{If our $m_w$ is higher than the actual weight of every rule in the best rule list, then the only approximation ratio we incur is that in the greedy approximation, which is a multiplicative factor of up to $1 - 1/e$. On the other hand, if our $m_w$ estimate is smaller than the actual maximum weight $m_w^*$ of a rule in the best list, then we may end up picking a sub-optimal rule in the best marginal rule function.}

\mrj{We now bound the sub-optimality of our chosen rule. Suppose the best rule to pick is $r^*$, and we pick rule $r$. Since we estimated $r^*$'s weight to be at most $m_w$, we would have estimated $r^*$ to have marginal value $\text{Est}(r^*) = \sum_{t \in r^*}(m_w - w(t))$ where $t \in r^*$ iterates over all tuples covered by $r^*$, and $w(t)$ is the weight of the current highest weight rule covering $t$. Since we picked $r$ over $r^*$, it means $\text{MarginalValue}(r) \geq \text{Est}(r^*)$. Suppose $m$ is the highest weight of a rule selected in the previous steps of BRS (so $w(t) \leq m$ for all $t$). Then the actual marginal value of $r^*$ is $\leq \frac{m_w - m}{m_w^* - m}\text{Est}(r^*)$. Thus our pick of $r$ is at most a $\frac{m_w - m}{m_w^* - m}$ factor worse than the optimal pick. If we always choose  our $m_w$ such we have $\frac{m_w - m}{m_w^* - m} \geq w$, for some constant $w < 1$, then our final score is guaranteed to be withing a $w(1-1/e)$ factor of the optimal score.}

@@

\techreporttext{
\vspace{-5pt}
\section{Extensions}\label{sec:extensions}
\techreporttext{
\subsection{Setting parameters $W$, $k$, $m_w$}\label{sec:setting-paramters}
Our system allows the user to tune the smart drill-down by adjusting a number of parameters. Having a lot of tunable parameters can increase the difficulty of using a system by increasing the decision-making burden on the user. To counteract this, we now provide ways to guide the user while selecting appropriate parameter values.

The $m_w$ parameter lets the user trade off the accuracy of the optimal rule-list and the running time. Ideally we want $m_w$ to equal the actual maximum weight of a rule in the optimal rule-list; this way we get full accuracy while also optimizing run-time. We cannot know the ideal value of $m_w$ in advance, but we can easily estimate it using sampling. We create a small random sample of tuples from the table, and run the BRS algorithm on it. Then the maximum weight $x$ of the output on the sample is likely to equal the maximum weight of the actual output. To account for sampling error, we can set $m_w$ to $2x$, which works well in practice.

\mrj{We now analyze a parametric family of weighting functions that generalizes our functions from Section~\ref{sec:weighting} and provides intuition for them. Let $W(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^k$, where the $w_c$'s and $k$ are parameters of the weighting function, and $o_{r,c}$ ($o$ stands for occurrence) is $1$ if $r(c) \neq \star$ and $0$ otherwise. The Size and Bits weighting functions can be seen as special cases of the above function with $k=1$ and with $w_c$ equal to $1$ for Size and $\text{log}(|c|)$ for Bits.}

\mrj{We want to estimate the Score contribution of a rule according to this weight function, which requires estimating the count of the rule. Suppose for each column $c$, the most frequent value occurs $f_c$ fraction of the time. If the probability of a value occurring in a column of a tuple is independent of other values in the tuple, the $Count$ of rule $r$ is approximately $|T|\Pi_{c \in C}f_c^{o_{r,c}}$. Thus the score of a rule set containing $r$ alone can be approximated by:
$$S(r) = \left(\sum_{c \in C} o_{r,c}w_c\right)^k|T|\Pi_{c \in C}f_c^{o_{r,c}}$$
}

\mrj{Now consider the rule that has the highest weight times count according to the above weighting function. We can approximate this rule by relaxing $o_{r,c}$ to a real number; that is by maximizing the estimated $S(r)$ subject to $0 \leq o_{r,c} \leq 1$. According to the KKT conditions, at the optimum, the partial derivative of $W(r)$ with respect to $o_{r,c}$ must be $\leq 0$ for each $c$ where $o_{r,c} = 0$, equal to $0$ if $0 < o_{r,c} < 1$ and $\geq 0$ otherwise. For each $c$, we have:
$$\frac{\partial S(r)}{\partial o_{r,c}} = S(r)(\frac{kw_c}{\sum_{c \in C} o_{r,c}w_c} + \ln f_c)$$}

\mrj{Thus if $x = \frac{k}{\sum_{c \in C} o_{r,c}w_c}$, then the maximum scoring rule should instantiate columns where $\frac{\ln f_c}{w_c} > x$ and some of the ones with $\frac{\ln f_c}{w_c} = x$. Thus for Size weighting, the highest score rule should select columns with high values of $\ln f_c$ i.e. highest frequency values. In contrast, Bits will select columns with highest values of $\log_{|c|}(f_c)$ i.e. values that are frequent relative to the number of distinct values in the column.}

\mrj{If we want the score to treat all columns equally, then the derivative must be $0$ for all $c$. This implies that for all $c$, we must have $w_c \propto \ln f_c$. Notice that Bits satisfies the above criterion if we assume that values in each column are uniformly distributed (that is, $f_c = 1/|c|$). This provides an intuitive justification for we might want to use the Bits weighting function.}

\mrj{We can also estimate the size of the optimal rule under the above weighting function. Specifically, the weighted fraction of columns with non-$\star$ values is given by $\frac{\sum_{c\in C} o_{r,c}w_c}{\sum_{c \in C} w_c}$ which equals $-k/(\sum_{c \in C} \ln f_c)$. Thus if we wanted our highest score rule to have $s$ fraction of the columns get instantiated, we can achieve that by setting $k = -s\sum_{c \in C} \ln f_c$.}

\mrj{The above reasoning also lets us estimate the $m_w$ parameter we need to use. Specifically, the highest score rule above has weight $(\frac{-\ln f_c}{kw_c})^k$.}

\mrj{We can use the above estimate to guess the size of the max scoring rule for the Bits weighting function. But note that in practice, values are not distributed uniformly (so $f_c$ is actually higher than $1/|c|$) and are not independent of each other, due to correlations. As a result, our estimate of the maximum size is a significant under-estimate. In our experiments (Section~\ref{sec:exp_qualitative}), we observe that use the Bits weighting function actually gives us rules of size $3$ on our datasets, while our estimate above is approximately $0.74$. As expected, the values displayed by the rules are ones that occur much more frequently relative to other values in that column.}
}
%This evaluates to $1.44$, while the observed value in our experiments is $10$.

\subsection{Dealing with Numerical Attributes}\label{sec:extensions-numerical}
\mrjdel{Our framework assumes that all attributes are categorical. Attributes that have a large domain tend to have fewer tuples per value, and hence don't appear in rule summaries. Thus our algorithm does not summarize information about numerical attributes.}

\mrjdel{However, we can modify the algorithm to deal with numerical attributes. Suppose we have a numerical attribute $A$. We can create buckets for values of $A$. We choose a number of buckets $b$, and divide the range of values of $A$ into $b$ intervals, with one bucket per interval. We can create buckets having equal range sizes, or decide their range such that there is an approximately equal number of tuples in each bucket. Then we run BRS, treating the bucket id as a categorical attribute. This is already done in our MD dataset, where numerical attributes like age are divided into buckets ($18-24$, $25-34$ and so on).}

\mrj{Our framework assumes that all attributes are categorical. Attributes that have a large domain tend to have fewer tuples per value, and hence don't appear in rule summaries. Thus our algorithm does not summarize information about numerical attributes. But it can easily be modified to do so by bucketizing a numerical attribute and treating the bucket id as a categorical attribute. This is already done in our MD dataset, where numerical attributes like age are divided into buckets ($18-24$, $25-34$ and so on).}

\vspace{-5pt}
\subsection{Using Sum instead of Count}\label{sec:extensions-sum}
\mrjdel{Throughout the paper, we define the total score of a rule-list using the marginal counts of rules in the list, and display the count of each rule in our table summary. However, if we have a numerical column (i.e. a `measure' column) in the table, it is straightforward to extend our summary to the `Sum' aggregate over that column instead. Suppose we are given a measure column $c_m$. Then the {\em Sum} for a rule can be defined to be the sum of $c_m$ values over all tuples covered by the rule. {\em MSum} of a rule $r$ in a rule-list $R$ is the sum of $c_m$ values over all tuples covered by $r$ and not covered by any rule in $R$ that occurs before $r$. The Score for $R$ becomes $\text{Score}(R) = \sum_{r\in R} MSum(r,R)W(r)$. Algorithm~\ref{algo:best-rule-set} can be modified to find the best rule set using the new definition of Score, simply by replacing $Count(r)$ by $Sum(r)$ and computing sum and marginal sum instead of count and marginal count in each pass over the table.}

\mrj{Throughout the paper, we define the total score of a rule-list using the marginal counts of rules in the list, and display the count of each rule in our table summary. However, if we have a numerical column (i.e. a `measure' column) in the table, it is straightforward to extend our summary to the `Sum' aggregate over that column instead. We can define $Sum$ and $MSum$ for a rule analogously to $Count$ and $MCount$, and $Score$ can be modified to use $MSum$ instead of $MCount$. We can then modify Algorithm~\ref{algo:best-rule-set} to maximize the new score by replacing $Count(r)$ by $Sum(r)$ and computing sum and marginal sum instead of count and marginal count in each pass over the table.}
}
